---
output: github_document
bibliography: inst/references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# did2s

<!-- badges: start -->
<!-- badges: end -->

The goal of did2s is to estimate TWFE models without running into the problem of staggered treatment adoption. 

For common issues, see this issue: [https://github.com/kylebutts/did2s/issues/12](https://github.com/kylebutts/did2s/issues/12)

## Installation

You can install did2s from CRAN with:

``` r
install.packages("did2s")
```

To install the development version, run the following:

```{r, eval = FALSE}
devtools::install_github("kylebutts/did2s")
```

## Two-stage Difference-in-differences [@Gardner_2021]

For details on the methodology, view this [vignette](http://kylebutts.com/did2s/articles/Two-Stage-Difference-in-Differences.html)

To view the documentation, type `?did2s` into the console.


The main function is `did2s` which estimates the two-stage did procedure. This function requires the following options:

- `yname`: the outcome variable
- `first_stage`: formula for first stage, can include fixed effects and covariates, but do not include treatment variable(s)!
- `second_stage`: This should be the treatment variable or in the case of event studies, treatment variables.
- `treatment`: This has to be the 0/1 treatment variable that marks when treatment turns on for a unit. If you suspect anticipation, see note above for accounting for this.
- `cluster_var`: Which variables to cluster on

Optional options:

- `weights`: Optional variable to run a weighted first- and second-stage regressions
- `bootstrap`: Should standard errors be bootstrapped instead? Default is False.
- `n_bootstraps`: How many clustered bootstraps to perform for standard errors. Default is 250.

did2s returns a list with two objects:

1. fixest estimate for the second stage with corrected standard errors.

### TWFE vs. Two-Stage DID Example

I will load example data from the package and plot the average outcome among the groups.

```{r load-data}

# Automatically loads fixest
library(did2s)

# Load Data from R package
data("df_het", package = "did2s")
```

Here is a plot of the average outcome variable for each of the groups:

```{r plot-df-het, fig.width=8, fig.height=4, fig.cap="Example data with heterogeneous treatment effects"}

# Mean for treatment group-year
agg <- aggregate(df_het$dep_var, by=list(g = df_het$g, year = df_het$year), FUN = mean)

agg$g <- as.character(agg$g)
agg$g <- ifelse(agg$g == "0", "Never Treated", agg$g)

never <- agg[agg$g == "Never Treated", ]
g1 <- agg[agg$g == "2000", ]
g2 <- agg[agg$g == "2010", ]


plot(0, 0, xlim = c(1990,2020), ylim = c(4,7.2), type = "n",
     main = "Data-generating Process", ylab = "Outcome", xlab = "Year")
abline(v = c(1999.5, 2009.5), lty = 2)
lines(never$year, never$x, col = "#8e549f", type = "b", pch = 15)
lines(g1$year, g1$x, col = "#497eb3", type = "b", pch = 17)
lines(g2$year, g2$x, col = "#d2382c", type = "b", pch = 16)
legend(x=1990, y=7.1, col = c("#8e549f", "#497eb3", "#d2382c"), 
       pch = c(15, 17, 16),
       legend = c("Never Treated", "2000", "2010"))


```


### Estimate Two-stage Difference-in-Differences 

First, lets estimate a static did. There are two things to note here. First, note that I can use `fixest::feols` formula including the `|` for specifying fixed effects and `fixest::i` for improved factor variable support. Second, note that `did2s` returns a `fixest` estimate object, so `fixest::etable`, `fixest::coefplot`, and `fixest::iplot` all work as expected.

```{r static}

# Static
static <- did2s(df_het, 
				yname = "dep_var", first_stage = ~ 0 | state + year, 
				second_stage = ~i(treat, ref=FALSE), treatment = "treat", 
				cluster_var = "state")

fixest::etable(static)

```

This is very close to the true treatment effect of ~2.23.

Then, let's estimate an event study did. Note that relative year has a value of `Inf` for never treated, so I put this as a reference in the second stage formula.

```{r event-study}

# Event Study
es <- did2s(df_het,
			yname = "dep_var", first_stage = ~ 0 | state + year, 
			second_stage = ~i(rel_year, ref=c(-1, Inf)), treatment = "treat", 
			cluster_var = "state")

```

And plot the results:

```{r plot-es, fig.cap="Event-study plot with example data"}

fixest::iplot(es, main = "Event study: Staggered treatment", xlab = "Relative time to treatment", col = "steelblue", ref.line = -0.5)

# Add the (mean) true effects
true_effects = head(tapply((df_het$te + df_het$te_dynamic), df_het$rel_year, mean), -1)
points(-20:20, true_effects, pch = 20, col = "black")

# Legend
legend(x=-20, y=3, col = c("steelblue", "black"), pch = c(20, 20), 
       legend = c("Two-stage estimate", "True effect"))

```


### Comparison to TWFE

```{r plot-compare, ig.cap="TWFE and Two-Stage estimates of Event-Study"}

twfe = feols(dep_var ~ i(rel_year, ref=c(-1, Inf)) | unit + year, data = df_het) 

fixest::iplot(list(es, twfe), sep = 0.2, ref.line = -0.5,
      col = c("steelblue", "#82b446"), pt.pch = c(20, 18), 
      xlab = "Relative time to treatment", 
      main = "Event study: Staggered treatment (comparison)")


# Legend
legend(x=-20, y=3, col = c("steelblue", "#82b446"), pch = c(20, 18), 
       legend = c("Two-stage estimate", "TWFE"))

```

# Citation

If you use this package to produce scientific or commercial publications, please cite according to:

```{r}
citation(package = "did2s")
```



# References



